---
title: "Course Project 1"
author: "Jared Lounsbery"
date: "October 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read the data


```{r read_data}
data <- read.csv("activity.csv", header = TRUE, na.strings = "NA")
data$date <- as.Date(data$date, format = "%Y-%m-%d")
```

## Number of steps per day


```{r numbsteps}
dailysteps <- tapply(data$steps, data$date, sum, na.rm=TRUE)
hist(dailysteps, xlab = "Number of steps per day")

dailymean <- mean(dailysteps, na.rm=TRUE)
dailymedian <- median(dailysteps, na.rm=TRUE)
```


The mean daily steps taken is `r dailymean`  

The median daily steps taken is `r dailymedian`  

##Average Number of Steps per 5-minute interval
```{r minutesteps}
minutesteps <- aggregate(data$steps, by=list(data$interval), mean, na.rm=TRUE)
plot(minutesteps, type="l", xlab = "5-Minute Interval", ylab = "Number of Steps")
maxavgsteps <- minutesteps$Group.1[max(minutesteps$x)]
```

The maximum average number of steps occurs in the `r maxavgsteps`th 5-minute of the day


##Imputing missing data

The number of missing data points is `r sum(is.na(data$steps))`

##Number of steps per day with imputed data
```{r imputing}
meansteps <- aggregate(data$steps, by=list(data$date), mean, na.rm=TRUE)
datanew <- data

for (i in 1:nrow(data)) {
    if (is.na(data$steps[i]) =="TRUE") {
       datanew$steps[i] = minutesteps$x[minutesteps$Group.1==datanew$interval[i]]
    }
}


imputedsteps <- tapply(datanew$steps, datanew$date, sum, na.rm=TRUE)
hist(imputedsteps, xlab = "Number of steps per day")

imputemeansteps <- mean(imputedsteps, na.rm=TRUE)
imputemediansteps <- median(imputedsteps, na.rm=TRUE)

```


The mean daily steps taken is `r imputemeansteps`  

The median daily steps taken is `r imputemediansteps`

As you can see, when we have replaced missing ('NA') data with imputed data based on time intervals, the mean and median daily steps increases.

```{r weekdays}
library(ggplot2)
library(gridExtra)

datanew$daytype <- weekdays(datanew$date)

datanew$daytype[datanew$daytype == "Monday" | 
                datanew$daytype == "Tuesday" | 
                datanew$daytype == "Wednesday"| 
                datanew$daytype == "Thursday"| 
                datanew$daytype == "Friday"] <- "weekday"

datanew$daytype[datanew$daytype == "Saturday" |
                datanew$daytype == "Sunday"] <- "weekend"


weekdaysteps <- aggregate(. ~ interval + daytype, data = datanew, mean, na.rm=TRUE)



plot1 <- with(weekdaysteps[weekdaysteps$daytype=="weekday",], ggplot(data=weekdaysteps,
                                                                     aes(x = interval, y = steps)) +
                                                              geom_line() +
                                                              ggtitle("weekdays"))
plot2 <- with(weekdaysteps[weekdaysteps$daytype=="weekend",], ggplot(data=weekdaysteps,
                                                                     aes(x = interval, y = steps)) +
                                                              geom_line()+
                                                              ggtitle("weekends"))

grid.arrange(plot1, plot2, nrow=2)
```
